---
title: "Portfolio"
author: "Bianca Rodas"
format: html
---

## Data Description

Link: https://github.com/rfordatascience/tidytuesday/blob/main/data/2021/2021-01-26/readme.md

Description: The dataset is gathered through organized plastic waste audits and clean-up events on a global scale. This dataset holds record of the date, location, brand name, types of plastic, and quantity.

Questions:
1.How does branded plastics compare to unbranded plastics collected?  
2.Which brands were most associated with the plastic found?
3.Which plastics types were more commonly found? Which is the least?


## Data Visualization

What do you want your final visualizations to look like?
> I want to have a bar graph comparing total counts of branded and unbranded plastics collected.
> Second visualization will be another bar graph with total counts of plastics for the top 10 brands found.
>Third visual could be a boxplot of different plastic types. 

What do you want to highlight on your final visualizations in order to answer your research questions? How do you plan to do that?
> I want to emphasize the top contributors to plastic pollution by filtering the top 10.

What is missing from your data or would need to change in your data to create these visualizations?
>I would need to use pivot to change the plastic types from columns to a new variable.I also need to group the companies and summarise the total counts.


## Data Cleaning

Do you need to reformat any variables into different types (e.g. factors, time, dates, strings)? Or remove information from variable values?
> no

Do you need to deal with any missing data, especially missing data coded other than NA?
> yes

Do you need to filter your data? How?
> yes, the top 10 highest counts of plastic collected by company

Do you need to create any new variables? What variables? How?
>no

Do you need to add new data (join) to your data? What data? How?
>no

Are there any variables you can exclude from your data?
>yes

Do you need pivot your data in any way? Why? How?
>yes, because the plastic types are column titles so I will use pivot_longer

Do you need to summarize any of the variables? Which ones? How?
>yes, the top 10 counts of plastics by company

What other aspects of your data need to be “fixed” in order to make your data visualizations?
>I don't think so.

```{r}
#| Label: Loading-libraries
#| echo: FALSE

library(tidyverse)
library(dplyr)
library(viridis)
```


```{r}
#| Label: Read-in-data
#| echo: FALSE

plastic <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-01-26/plastics.csv", na = c("NA", "n/a", "unknown", "null", "NULL"))

```


```{r}
#| Label: Clean-data
#| echo: FALSE

plastic <- plastic |> 
  mutate(brand= if_else(parent_company %in% c("Unbranded") | is.na(parent_company), "Unbranded", "Branded"))
```


```{r}
top_companies <- plastic |> 
  filter(!parent_company %in% c("Grand Total", "Unbranded") & !is.na(parent_company)) |> 
  group_by(parent_company) |> 
  summarise(sum_grand_total = sum(grand_total)) |> 
  arrange(desc(sum_grand_total)) |> 
  slice_head(n = 10)
```

## Visualizations

```{r}
ggplot(plastic, aes(x=brand, y=grand_total, fill=brand))+
  geom_col()+
  theme_bw()+
  theme(legend.position = "none")+
  scale_fill_viridis(discrete=T,
                     option = "plasma",
                     begin=0.2, 
                     end=0.9)+
  labs(x="",
       y= "Plastic Counts",
       title ="Plastic Found with an Associated Brand")
```


```{r}
ggplot(top_companies, aes(x=sum_grand_total, 
                              y=reorder(parent_company, (sum_grand_total))
                              , fill= parent_company))+
  geom_col(color="black",size=0.4)+
  theme_bw()+
  scale_fill_viridis_d(option = "plasma", guide = "colourbar")+
  labs(title="Companies identified during plastic collection",
       y= "",
       x="Total Trash Count")
```




